---
title: "Urbanization Impacts on Evapotranspiration Across Various Spatio-temporal Scales"
author: "Mazrooei et al"
output: html_document
---


```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(knitr)
#opts_chunk$set(fig.width=12, fig.height=8)
knitr::opts_chunk$set(echo = FALSE)

library(raster)
library(dplyr)
library(plotly)
library(leaflet)

```

```{r, echo=FALSE, include=FALSE, message=FALSE, warning=FALSE}
## FIG 1
# Path_usefulGIS = "G:/My Drive/NCSU-laptop sharing/Research/GIS/useful shapefiles"
# Path_ETdata = "G:/My Drive/NCSU-laptop sharing/Research/POWELL CENTER/Data/ET_LC/ET_GLOBAL_Worldsubdivs_Koppen/"
# filename = paste0(Path_ETdata,"MatlabOutput_Global_ET_Annual_mmperyear.txt")
# filename_2 = paste0(Path_ETdata,"To_Join_Koppen_Codes_Names.txt")
# 
# system.time({ layer_org <- shapefile(paste0(Path_usefulGIS,"/Worldsubdivs_Koppen_simplified.shp")) })
# system.time({ layer_conus_org <- shapefile(paste0(Path_usefulGIS,"/US_Climate_Divisions_GCS_NAD_1983_addedFID1.shp")) })
# layer <- layer_org
# layer_conus <- layer_conus_org
# layer$FID <- layer$FID_1
# layer_conus$FID <- layer_conus$FID_1
# #layer@data$id <- rownames(layer@data)
# 
# DeltaET.data <- read.csv(filename, header=TRUE, sep=",")
# #DeltaET.data$DeltaET_Annual_mmperyear = -DeltaET.data$DeltaET_Annual_mmperyear #\DeltaET = ET_{urban} - ET_{non-urban}
# layer <- merge(layer, DeltaET.data, by="FID")
# Koppen_classes <- read.csv(filename_2 , header=TRUE, sep="\t")
# layer <- merge(layer, Koppen_classes, by="FID")
# 
# filename_ET_conus = paste0("G:/My Drive/NCSU-laptop sharing/Research/POWELL CENTER/Data/ET_LC/ET_US_ClimateDivs/","MatlabOutput_ClimateDivs_ET_Annual_mmperyear.txt")
# DeltaET.data_conus <- read.csv(filename_ET_conus, header=TRUE, sep=",")
# layer_conus <- merge(layer_conus, DeltaET.data_conus, by='FID')
# 
# 
# ## FIG 2
# ETdata <- read.delim(paste0(Path_ETdata,"./MatlabOutput_Global_Lambda_LogAI.txt"), header = TRUE, sep = "\t")
# ETdata$FID <- (1:dim(ETdata)[1])-1
# ETdata$log_bar_to_veg <- as.numeric(ETdata$log.A_barren.A_veg.)
# ETdata <- ETdata[,c("FID","log.AI.","Lambda_ET","log_bar_to_veg")]
# 
# ind <- rowSums(is.na(ETdata[,2:4])) == 0 
# ETdata <- ETdata[ind,]
# 
# ETdata <- merge(ETdata, Koppen_classes, by="FID")
# 
# #Path_usefulGIS = "G:/My Drive/NCSU-laptop sharing/Research/GIS/useful shapefiles"
# #layer = shapefile(paste0(Path_usefulGIS, "/Worldsubdivs_Koppen_simplified.shp"))
# #meta_KCC <- as.data.frame(layer)
# #saveRDS(meta_KCC,file = "./meta_KCC.rds")
# meta_KCC <- readRDS(paste0(Path_ETdata,"/meta_KCC.rds"))
# meta_KCC$FID <- as.numeric(meta_KCC$FID_1)
# DF <- merge(ETdata, meta_KCC, by="FID")
# save(list = ls(), file = "data.Rdata")

load("data.Rdata")

```

## Abstract {.tabset}
Urbanization has been shown to locally increase the nighttime temperatures creating urban heat islands, which partly arise due to evapotranspiration (ET) reduction. It is unclear how the direction and magnitude of the change in local ET due to urbanization varies globally across different climatic regimes. This knowledge gap is critical, both for the key role of ET in the energy and water balance accounting for the majority of local precipitation, and for reducing the urban heat island effect. We explore and assess the impacts of urbanization on monthly and mean annual ET across a range of landscapes from local to global spatial scales. Remotely sensed land cover and ET available at 1 km resolution are used to quantify the differences in ET between urban and surrounding non-urban areas across the globe. The observed patterns show the statistically significant difference between urban and non-urban ET can be estimated to first order as a function of local hydroclimate with arid regions seeing increased ET due to urbanization, and humid regions seeing decreased ET. Cities under cold climates also evaporate more than its non-urban surroundings during the winter as the urban micro-climate has increased energy availability resulting from human activity.  Increased ET in arid cities clearly arises from municipal water withdrawals and increased irrigation during drought conditions further increases the ET from arid urban cities compared to non-urban ET. These results can help inform planners to improve the integration of environmental conditions into the design and management of urban landscapes.

<center> $\Delta ET = ET_{urban} - ET_{non-urban}$ </center>
<center> $\lambda = ET_{urban} / ET_{non-urban}$ </center>

### $\Delta$ET Map {.tabset}

#### Global
```{r, echo = FALSE, message=FALSE, warning=FALSE}

bins <- c(-9999,seq(-400,400,100),9999)
pal <- colorBin("RdBu", domain = layer@data$DeltaET_Annual_mmperyear, bins = bins, reverse = TRUE)

labels <- sprintf("FID: %s<br/><strong>Country:</strong> %s<br/><strong>Main Climate class:</strong> %s<br/><strong>Precipitation class:</strong> %s<br/><strong>Temperature class:</strong> %s<br/><strong>urban ET:</strong> %g [mm/year]<br/><strong>non-urban ET:</strong> %g [mm/year]<br/><strong>%s</strong> %g [mm/year]", layer@data$FID, layer@data$NAME, layer@data$Koppen_main_climate, layer@data$Koppen_precipitation, layer@data$Koppen_temperature, layer@data$urban_ET_Annual_mmperyear, layer@data$nonurban_ET_Annual_mmperyear, "&Delta;ET=", layer@data$DeltaET_Annual_mmperyear ) %>% lapply(htmltools::HTML)

leaflet(layer, width = "100%") %>% addProviderTiles("CartoDB.DarkMatter") %>% setView(lng = 0, lat = 0, zoom = 2) %>%
  addPolygons(fillColor = ~pal(DeltaET_Annual_mmperyear), color = "gray", weight = 0.6, smoothFactor = 0.5, opacity = 1, dashArray = "3", fillOpacity = .71,
    highlightOptions = highlightOptions(color = "white", weight = 2,
    bringToFront = TRUE), label = labels, popup =labels ,
    labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "2px 4px"),
    textsize = "15px", direction = "top")) %>%
  addLegend(pal = pal, values = ~layer@data$DeltaET_Annual_mmperyear, opacity = .71, title = "&Delta;ET [mm/year]", position = "bottomright")

```

#### CONUS
```{r, echo = FALSE, message=FALSE, warning=FALSE}
bins <- c(-9999,seq(-400,400,100),9999)
pal <- colorBin("RdBu", domain = layer_conus@data$DeltaET_Annual_mmperyear, bins = bins, reverse = TRUE)

labels <- sprintf("FID: %s<br/><strong>State:</strong> %s<br/><strong>Name:</strong> %s<br/><strong>urban ET:</strong> %g [mm/year]<br/><strong>non-urban ET:</strong> %g [mm/year]<br/><strong>%s</strong> %g [mm/year]", layer_conus@data$FID, layer_conus@data$ST, layer_conus@data$NAME, layer_conus@data$urban_ET_Annual_mmperyear, layer_conus@data$nonurban_ET_Annual_mmperyear, "&Delta;ET=", layer_conus@data$DeltaET_Annual_mmperyear ) %>% lapply(htmltools::HTML)

leaflet(layer_conus, width = "100%") %>% addProviderTiles("CartoDB.DarkMatter") %>% setView(lng = -95, lat = 37.5, zoom = 4) %>%
  addPolygons(fillColor = ~pal(DeltaET_Annual_mmperyear), color = "gray", weight = 0.6, smoothFactor = 0.5, opacity = 1, dashArray = "3", fillOpacity = .71,
    highlightOptions = highlightOptions(color = "white", weight = 2,
    bringToFront = TRUE), label = labels, popup =labels ,
    labelOptions = labelOptions(style = list("font-weight" = "normal", padding = "2px 4px"),
    textsize = "15px", direction = "top")) %>%
  addLegend(pal = pal, values = ~layer_conus@data$DeltaET_Annual_mmperyear, opacity = .71, title = "&Delta;ET [mm/year]", position = "bottomright")

```


### $\lambda$ vs. log(AI) {.tabset}
#### v1 {.smaller}
```{r, echo = FALSE, message=FALSE, warning=FALSE}
cmin <- -12
cmax <- 2
DF$Mycolor <- DF$log_bar_to_veg; DF$Mycolor[DF$Mycolor>cmax] <- cmax; DF$Mycolor[DF$Mycolor<cmin] <- cmin;


plot_ly(DF, x = ~log.AI., y = ~Lambda_ET, 
        type = 'scatter',  mode='markers', 
        marker=list(size=4, color=~Mycolor, colorscale='Jet', reversescale=F,
                    colorbar=list(title='Ratio log of Area \n barren/vegetated')),
        text = sprintf("FID: %d\nCountry: %s\nMain Climate class: %s\nPrecipitation class: %s\nTemperature class: %s\n log(bar/veg): %.2f", DF$FID, DF$NAME, DF$Koppen_main_climate, DF$Koppen_precipitation, DF$Koppen_temperature, DF$log_bar_to_veg),
        hovertemplate = paste(
          "<b>%{text}<br>",
          "%{yaxis.title.text}: %{y:.2f}<br>",
          "%{xaxis.title.text}: %{x:.2f}",
          "<extra></extra>"
        )
) %>%
layout(xaxis = list(range = c(-1,5)), yaxis = list(range = c(0, 3)))
```

#### v2
```{r, echo = FALSE, message=FALSE, warning=FALSE}
plot_ly(DF, x = ~log.AI., y = ~Lambda_ET, color=~Koppen_main_climate, 
        colors=c("red","dodgerblue","gray","cyan","blueviolet","greenyellow"),
        type = 'scatter',  mode='markers', 
        marker=list(size=4),
        text = sprintf("FID: %d\nCountry: %s\nMain Climate class: %s\nPrecipitation class: %s\nTemperature class: %s\n log(bar/veg): %.2f", DF$FID, DF$NAME, DF$Koppen_main_climate, DF$Koppen_precipitation, DF$Koppen_temperature, DF$log_bar_to_veg),
        hovertemplate = paste(
          "<b>%{text}<br>",
          "%{yaxis.title.text}: %{y:.2f}<br>",
          "%{xaxis.title.text}: %{x:.2f}",
          "<extra></extra>"
        )
) %>%
  layout(xaxis = list(range = c(-1,5)), yaxis = list(range = c(0, 3))) %>%
    colorbar(title='Main Climate')
```


